Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-09-29 manuscript/manuscript.tex
G's changes to disco

Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
that diversity is reduced at linked neutral variation as an advantageous
polymorphism sweeps to fixation, over four decades of theoretical and empirical
research has enriched our understanding of linked selection. This work has led
to the discovery of selected genes from genomic signals of hitchhiking
\parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of linked
selection
\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
and background selection \parencite{Charlesworth1993-gb} models that attempt
explain this pattern. Since, much attention has focused on quantifying the
impact of background selection \parencite{McVicker2009-ax} and developing
theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
reductions in diversity into those caused by full and partial sweeps, and those
caused by background selection \parencite{Elyashiv2016-vt}. However, as other
models of soft sweeps from standing variation emerged
\parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
difficult endeavor \parencite{Przeworski2005-bg} addressed using new coalescent
models \parencite{Berg2015-xj} and machine learning methods
\parencite{Schrider2017-yx}. An alternative approach to understanding the
genome-wide effects of selection on standing variation, e.g. selection on an
infinitesimal polygenic trait, stems from an early quantitative genetic model
of linked selection \parencite{Robertson1961-ho} and its later developments
(\cite{Santiago1995-hx,Santiago1998-bs}; see also \cite{Barton2000-zg} for a
comparison of these models with classic hitchhiking models). Implicit in these
models is that autocovariance between allele frequency change is created when
there is heritable fitness variation in the population, a signal that can be
readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
we provide the first empirical evidence of linked selection acting on standing
variation over very short timescales.

@@

While we must keep in mind that the studies analyzed here were all laboratory
populations and natural selection in the wild will likely differ, overall we
have found that linked selection has a remarkably strong effect for such short
study durations. Depending on how many loci affect fitness, such a strong
effect of linked selection may not be differentiable from genetic drift using
only single contemporary population samples. In this way, temporal data allows
us to sidestep the key problem of detecting selection from standing variation:
that the genomic footprint leaves too soft of a signature to differentiate from
a background of genetic drift. In fact we find that the temporal covariance
signal is detectable even in the most extremely difficult to detect soft sweep
case: polygenic selection on an infinitesimal trait. In our reanalysis of the
\textcite{Castro2019-uk} data, by showing the covariance signal remains even
after excluding the chromosomes containing two large-effect loci, we showed the
covariance signal we detect is indeed driven by polygenic fitness variation
(and confirm a point in the original study that tibiae length, aside from these
two large effect loci, is a highly polygenic trait).

>>>>

Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
that diversity is reduced at linked neutral variation as an advantageous
polymorphism sweeps to fixation, over four decades of theoretical and empirical
research has enriched our understanding of linked selection.

An alternative approach to understanding the
genome-wide effects of selection on standing variation, e.g. selection on an
infinitesimal polygenic trait, stems from an early quantitative genetic model
of linked selection \parencite{Robertson1961-ho} and its later developments
(\cite{Santiago1995-hx,Santiago1998-bs}; see also \cite{Barton2000-zg} for a
comparison of these models with classic hitchhiking models). Implicit in these
models is that autocovariance between allele frequency change is created when
there is heritable fitness variation in the population, a signal that can be
readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
we provide the first empirical evidence of linked selection acting on standing
variation over very short timescales.

Depending on how many loci affect fitness, such a strong
effect of linked selection may not be differentiable from genetic drift using
only single contemporary population samples. In this way, temporal data allows
us to sidestep the key problem of detecting selection from standing variation:
that the genomic footprint leaves too soft of a signature to differentiate from
a background of genetic drift. In fact we find that the temporal covariance
signal is detectable even in the most extremely difficult to detect soft sweep
case: polygenic selection on an infinitesimal trait.

%In our reanalysis of the
%\textcite{Castro2019-uk} data, by showing the covariance signal remains even
%after excluding the chromosomes containing two large-effect loci, we showed the
%covariance signal we detect is indeed driven by polygenic fitness variation
%(and confirm a point in the original study that tibiae length, aside from these
%two large effect loci, is a highly polygenic trait).

%This work has led
%to the discovery of selected genes from genomic signals of hitchhiking
%\parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of linked
%selection
%\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
%leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
%and background selection \parencite{Charlesworth1993-gb} models that attempt
%explain this pattern. Since, much attention has focused on quantifying the
%impact of background selection \parencite{McVicker2009-ax} and developing
% theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
% reductions in diversity into those caused by full and partial sweeps, and those
% caused by background selection \parencite{Elyashiv2016-vt}. However, as other
% models of soft sweeps from standing variation emerged
% \parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
% detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
% difficult endeavor \parencite{Przeworski2005-bg} addressed using new coalescent
% models \parencite{Berg2015-xj} and machine learning methods
%\parencite{Schrider2017-yx}.

@@

While we must keep in mind that the studies analyzed here were all laboratory
populations and natural selection in the wild will likely differ, overall we
have found that linked selection has a remarkably strong effect for such short
study durations.
